---
title: "PA1_template"
author: "Pavel Pustovoyt"
date: "22 January 2016"
output: html_document
---

Required libraries: `ggplot2, dplyr`

### Loading and preprocessing the data

Loading `activity` data:

```{r}
options(scipen=100, digits=1)
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
activity <- read.csv("../data/activity.csv", stringsAsFactors = FALSE)
```


### What is mean total number of steps taken per day?

First lets remove `NA` values and group data by `date` summing number of `steps`:
```{r, echo=TRUE}
activityNoNA <- filter(activity, !is.na(steps))
sumByDay <- summarise(group_by(activityNoNA, date), steps = sum(steps))
head(sumByDay)
```

Now lets plot histogram:
```{r, echo=TRUE}
ggplot(sumByDay) + geom_histogram(aes(steps, fill=..count..), binwidth = 2000) + scale_fill_gradient(low="skyblue", high = "skyblue4")
```

Average steps per day:
```{r, echo=TRUE}
mean(sumByDay$steps) #sum(activity$steps, na.rm =TRUE)/length(unique(activity$date))
```
Median:
```{r, echo=TRUE}
median(sumByDay$steps)
```



### What is the average daily activity pattern?

Grouping data by `interval` and taking `mean`:
```{r, echo=TRUE}
avgByInterval <- summarise(group_by(activityNoNA, interval), steps = mean(steps))
head(avgByInterval)
```

Time series plot of the average number of steps taken:
```{r, echo=TRUE}
ggplot(avgByInterval) + geom_line(aes(x = interval, y = steps, colour=steps)) + scale_colour_gradient(low="skyblue", high = "darkblue")
```

The 5-minute interval that, on average, contains the maximum number of steps:
```{r, echo=TRUE}
max(avgByInterval$steps)
```


### Imputing missing values

Total number of missing values:
```{r, echo=TRUE}
sum(is.na(activity$steps))
```

Filling missing interval data with average for that interval across all days:

```{r, echo=TRUE}
#Joining average value for every interval
activityFilled <- merge(activity, avgByInterval, by = c("interval"))
#Ordering by date and interval
activityFilled <- arrange(activityFilled, date, interval)
#Filling empty values
activityFilled$steps <- ifelse(is.na(activityFilled$steps.x), activityFilled$steps.y, activityFilled$steps.x)
activityFilled <- select(activityFilled, -steps.x, -steps.y)
```

### Imputing missing values

Histogram of average steps per day after filling `NA` values:
```{r, echo=TRUE}
sumByDayFilled <- summarise(group_by(activityFilled, date), steps = sum(steps))
ggplot(sumByDayFilled) + geom_histogram(aes(steps, fill=..count..), binwidth = 2000) + scale_fill_gradient(low="skyblue", high = "skyblue4")
```

Mean after filling values:
```{r, echo=TRUE}
mean(sumByDayFilled$steps)
```

Median after filling values:
```{r, echo=TRUE}
median(sumByDayFilled$steps)
```

### Are there differences in activity patterns between weekdays and weekends?

Defining day of week and type of day for every interval:
```{r, echo=TRUE}
activityFilled$day <- weekdays(as.Date(activityFilled$date))
activityFilled$dayType <- as.factor(ifelse(activityFilled$day %in% c("Sunday","Saturday"), "weekend","weekday"))
head(activityFilled)
```

Gouping by `interval` and `dayType`:
```{r, echo=TRUE}
avgByIntervalFilled <- summarise(group_by(activityFilled, interval, dayType), steps = mean(steps))
head(avgByIntervalFilled)
```

Plotting graph to compare weekends and weekdays(to make it two-paneled you just need to uncomment end of line, personally I like both lines on the same plot):
```{r, echo=TRUE}
ggplot(avgByIntervalFilled) + geom_line(aes(x = interval, y = steps, group=dayType, colour=dayType)) # + facet_grid(dayType~.)
```

Looks like guy is very active in the mornings =)
